Deep Probabilistic Models for Camera Geo-Calibration

### 2018-08-20 chapters/introduction/introduction.tex
a check from the abstract to main contributions

\begin{chapquote}{Alan Turing}
  ``I propose to consider the question: Can machines think?''
\end{chapquote}
The ultimate goal of image understanding is to transfer visual
signals (images/videos) into abstract symbolic descriptions of the
world which are helpful for decision making.
However, understanding images is a not trivial task for machines.
In order to handle various complicated scenarios, researchers divide
image understanding into different computer vision tasks: pedestrian
detection for autonomous driving, face recognition for security
system, and image retrieval for image searching engines.

@@

\subsection{Geo-Calibration of Outdoor Images}
Automatic image geo-calibration of outdoor images continues to grow in
importance as a direct result of the increasing amount of imagery
available via the Internet.  Solving this problem is of great value
for a wide variety of fields, with potential applications ranging from
the forensic sciences~\cite{stylianou13jane} to environmental
monitoring~\cite{zhang2012mining}.  Conceptually the geo-calibration
task includes: identifying the camera pose ({\em pose estimation}),
the camera geographic location ({\em geo-localization}), and the time
the query image was captured ({\em time estimation}).

@@

\subsection{Deep Learning in Geo-Calibration}
In recent years, deep neural networks (DNNs) were proven to be
extremely successful in many computer vision areas. 
It is widely accepted that DNNs have excellent performance in
high-level visual feature learning.
This valuable ability provides researchers powerful tools to
solve the challenging geo-calibration tasks.
Weyand and Kostrikov \etal~\cite{planet} propose a convolutional
network to predict the geographic location of the input
image. Walch \etal~\cite{walch2017image} aggregate learned CNN
features with LSTM to generate global image representation for camera
localization.  In pose estimation aera, Work~\cite{zhai2016horizon,
workman2016horizon, hold2017perceptual} has been done to to identify
the horizon line with CNNs (one can derive the camera roll and pitch
angles from the horizon line position giving the camera focal length).

Closely related to camera geo-calibration, studies also have been done
to identify the camera intrinsics using deep neural networks.
Workman \etal~\cite{workman2015deepfocal} develop a CNN to estimate the camera
focal length. Kendall \etal~\cite{kendall2015convolutional} propose a
neural network that can identify 6-DOF camera parameters, but their
method only limited to specific scenes.

\subsection{Ground-to-Aerial Geo-Calibration}
Data-driven scene matching approaches are widely used to for
geo-calibration.
When localizing a camera, data-driven algorithms search for k-nearest
neighbors of the query image in the database of the geo-tagged imagery.
To reduce the work of querying, the searching process is usually
carried out in some feature space with applicable number of  
dimensions~\cite{im2gps, li2010location,zamir2010accurate}.
However, as items found in most reference dataset, geo-tagged
ground-level images collected from the internet are usually noisy and
sparse in geographic space. To achieve more accurate geo-calibration
results, we need alternative image resources with more accurate
geographic tags and denser spatial coverage.

Benefited from the fast growing monitoring satellite and drone markets,
the geo-tagged aerial imagery becomes more publicly accessible
than ever before. Aerial imagery downloading services like
Microsoft Bing Map provide aerial images of various
resolutions, which can cover most areas of the world with accurate
geographic tags, making aerial imagery a rich source for
reference database. We refer to the geo-calibration methods
which use the aerial imagery as geographic reference as
ground-to-aerial geo-calibration.

Recent work on ground-to-aerial
geo-calibration~\cite{lin2013cross,lin2015learning,workman2015geocnn,workman2015wide}
has shown that convolutional neural networks are capable of extracting
features from geo-tagged aerial imagery that can be matched to
features extracted from the ground imagery.  Vo
\etal~\cite{vo2016localizing} extend this line of work, demonstrating
improved geo-localization performance by applying an auxiliary loss
function to regress the ground-level camera orientation with respect
to the aerial image. Hu and Feng \etal~\cite{mh2018cvm} achieve the
state-of-the-art performance in geo-localization by extracting
cross-domain global features using the
NetVLAD~\cite{arandjelovic2016netvlad} technique.

@@

Automatic feature engineering has drawn a lot of attention over the
past years. Compared to the manual feature engineering,
automatic feature engineering approaches are able to to learn useful
features without strong expert knowledge about the data. Among these
approaches, deep neural networks are wildly accepted by researchers as
good tools to learn high-level image representations.  The most
commonly used approach for deep feature learning is full supervision,
which usually requires manually annotated
data~\cite{yosinski2014transferable,zhou2016learning,wen2016discriminative}.
Unfortunately, such data is not always available for certain tasks.
In contrast, unlabeled data is offen plentiful (\ie cellphone photos
in Instagram or Youtube videos).

To exploit these unlabeled data, recent work has explored
self-supervision methods (sometimes referred to as unsupervised learning or
pretext tasks) for training deep neural networks that capture useful
visual representations~\cite{doersch2015unsupervised,pathak2016context}. 
These methods typically
exploit some known quantity of the data (e.g., pixel color values) to
avoid expensive manual annotation. As a byproduct, a useful visual
representation is learned.
For example, Zhang et al.~\cite{zhang2016colorful} show how image
colorization (synthesizing colors for a grayscale image) is a powerful
pretext task for learning visual representations. Pathak et
al.~\cite{pathak2017learning} exploit low-level motion-based grouping
cues for unsupervised feature learning.  
%
As new learning technique address the lack of massive amounts of
labeled data, domain adaptation adapt the feature generating model to
another feature domain so that it improve the performance when dealing
with new input
data~\cite{fernando2013unsupervised,fernando2015joint,saenko2010adapting,wang2016actions,tinghui2016flow}.
Duan \etal~\cite{duan2012learning} proposed to learn a linear
projection to transfer from the source feature subspace to the target
subspace. Sohn \etal~\cite{sohn2017unsupervised} improved the
performance of video face recognition using an adversarial-based
approach to adapt the network to unlabled video imagery.

@@

We make contributions in three main areas: 1) We proposed a general
probabilistic model to jointly estimate the geo-calibration
parameters, 2) we decompose the full geo-calibration into several
partial geo-calibration tasks, each of which provides a strong
constraint function that fits in the general model, and 3) we show
that learning to geo-calibrate a camera is helpful for high-level
image feature extraction.

The first two contributions consist of our main approach. Our
algorithm feeds inputs (images, time and manual annotations) to
different constraint models, which work as constraint functions. Then
we fuse these constraint functions to simulate the joint distribution
over the complete geo-calibration parameters. The flowchart of
our approaches refers to \figref{intro:architecture}.

Furthermore, we also explore to learning high-level
visual features and geo-calibrating cameras simultaneously.
We believe that the geo-calibration parameters are closely related to
the image representation. 
Imagine the process of a human doing geo-localization, one could
roughly guess where a photo was taken when she sees an polar bear on
an iceberg on it. In this process, learning high-level features
becomes a necessary step of abstracting visual hints to localize
the camera.
Based on this belief, we designed network to learn useful visual
features by exploiting geographic information, which provides new ways
for automatic feature learning.

>>>>

\begin{chapquote}{Morgan Freeman in {\em Transcendence (2014 film)}
\\ \todo{don't laugh.}}
  ``Geo-locating suspects in real time? I've never seen anything like
it.''
\end{chapquote}

The ultimate goal of image understanding is to transfer visual
signals (images/videos) into abstract symbolic descriptions of the
world which are helpful for decision making.
However, understanding images is a not trivial task for machines.
In order to handle various complicated scenarios, researchers divide
image understanding into different computer vision tasks: pedestrian
detection for autonomous driving, face recognition for security
system, and image retrieval for image searching engines.

@@

\subsection{Geo-Calibration of Outdoor Images}
Automatic image geo-calibration of outdoor images continues to grow in
importance as a direct result of the increasing amount of imagery
available via the Internet.  Solving this problem is of great value
for a wide variety of fields, with potential applications ranging from
the forensic sciences~\cite{stylianou13jane} to environmental
monitoring~\cite{zhang2012mining}.  Conceptually the geo-calibration
task includes: identifying the camera pose ({\em pose estimation}),
the camera geographic location ({\em geo-localization}), and the time
when the query image was captured ({\em time estimation}).

@@

\subsection{Deep Learning in Geo-Calibration}
In recent years, deep neural networks (DNNs) were proven to be
extremely successful in many computer vision areas. 
It is widely accepted that DNNs have excellent performance in
high-level visual feature learning.
This valuable ability provides researchers powerful tools to
solve the challenging geo-calibration tasks.
Weyand and Kostrikov \etal~\cite{planet} propose a convolutional
network to predict the geographic location of the input
image. Walch \etal~\cite{walch2017image} aggregate learned CNN
features with LSTM to generate global image representation for camera
localization.  In pose estimation aera, Work~\cite{zhai2016horizon,
workman2016horizon, hold2017perceptual} has been done to identify
the horizon line with CNNs (one can derive the camera roll and pitch
angles from the horizon line position giving the camera focal length).

Closely related to camera geo-calibration, problems of identifying
camera intrinsics using deep neural networks have also been studied.
Workman \etal~\cite{workman2015deepfocal} develop a CNN to estimate the camera
focal length. Kendall \etal~\cite{kendall2015convolutional} propose a
neural network that can identify 6-DOF camera parameters for specific
scenes.

\subsection{Ground-to-Aerial Geo-Calibration}
Data-driven scene matching approaches are commonly used in
geo-calibration.
When localizing a camera, data-driven algorithms search for k-nearest
neighbors of the query image in a database of the geo-tagged imagery.
To reduce the work of querying, the searching process is usually
carried out in feature space of applicable number of  
dimensions~\cite{im2gps, li2010location,zamir2010accurate}.
However, due to limited accuracies of GPS signals and the biased
distribution of human population, geo-tagged ground-level images
collected from smart devices are usually noisy and sparse in
geographic space. To achieve better geo-calibration results, we
need alternative image resources with more accurate geographic tags
and more complete spatial coverage.

Benefited from the fast growing monitoring satellite and drone markets,
the geo-tagged aerial imagery becomes more publicly accessible
than ever. Aerial imagery downloading services like
Microsoft Bing Map provide aerial images of various
resolutions, which cover most areas of the world with accurate
geographic registration, making aerial imagery a rich source for
reference databases. We refer to geo-calibration
methods which use the aerial imagery for geographic reference as
ground-to-aerial geo-calibration.

Recent work on ground-to-aerial
geo-calibration~\cite{lin2013cross,lin2015learning,workman2015geocnn,workman2015wide}
has shown that convolutional neural networks are efficient for extracting
features from geo-tagged aerial imagery that can be matched to
features extracted from the ground imagery.  Vo
\etal~\cite{vo2016localizing} extend this line of work, demonstrating
improved geo-localization performance by applying an auxiliary loss
function to regress the ground-level camera orientation with respect
to the aerial image. Hu and Feng \etal~\cite{mh2018cvm} achieve the
state-of-the-art performance in geo-localization by extracting
cross-domain global features using the
NetVLAD~\cite{arandjelovic2016netvlad} technique.

@@

Automatic feature engineering has drawn a lot of attention over the
past years. Compared to the manual feature engineering,
automatic feature engineering approaches allow to learn useful
features without strong expert knowledge about the data. Among these
approaches, deep neural networks are wildly accepted as
good tools to learn high-level image representations.  The most
commonly used approach for deep feature learning is full supervision,
which usually requires manual
annotations~\cite{yosinski2015transferable,zhou2016learning,wen2016discriminative}.
Unfortunately, such data is not always available for certain tasks.
In contrast, unlabeled data is often plentiful (\ie cellphone photos
in Instagram or Youtube videos).

To exploit these unlabeled data, recent work has explored
self-supervision methods (sometimes referred to as unsupervised learning or
pretext tasks) for training deep neural networks that capture useful
visual representations~\cite{doersch2015unsupervised,pathak2016context}. 
These methods typically
exploit some known quantity of the data (e.g., pixel color values) to
avoid expensive manual annotation. As a byproduct, a useful visual
representation is learned.
For example, Zhang et al.~\cite{zhang2016colorful} show how image
colorization (synthesizing colors for a grayscale image) is a powerful
pretext task for learning visual representations. Pathak et
al.~\cite{pathak2017learning} exploit low-level motion-based grouping
cues for unsupervised feature learning.  
%
As new learning technique that addresses the lack of massive amounts
of labeled data, domain adaptation forces the feature generating model
to adapt another feature domain so that it improves the performance
when dealing with new input
data~\cite{fernando2013unsupervised,fernando2015joint,saenko2010adapting,wang2016actions,tinghui2016flow}.
Duan \etal~\cite{duan2012learning} propose to learn a linear
projection to transfer from the source feature subspace to the target
subspace. Sohn \etal~\cite{sohn2017unsupervised} improve the
performance of video face recognition using an adversarial-based
approach to adapt the network to unlabeled video imagery.

@@

We make contributions in three main areas: 1) we proposed a general
probabilistic model to jointly estimate the geo-calibration
parameters, 2) we decomposed the full geo-calibration into several
partial geo-calibration tasks, each of which provides a strong
constraint function that fits into the general model, and 3) we show
that learning to geo-calibrate a camera is helpful for high-level
image feature extraction.

The first two contributions form our main approach. Our
algorithm feeds inputs (images, time and manual annotations) to
different constraint models, which work as constraint functions. Then
we fuse these constraint functions to simulate the joint distribution
over the complete geo-calibration parameters. We present the flowchart of
our approach in \figref{intro:architecture}.

Furthermore, we also explore to learning high-level visual features
and geo-calibrating cameras simultaneously.  We believe that camera
poses, geographic locations and image capture times are closely
related to image representations. 
Imagine a human geo-localizing a camera from an image it took, one
could roughly guess where a photo was taken when she sees an polar
bear on an iceberg. In this scenario, learning high-level
features becomes a necessary process of abstracting visual hints to
localize the camera.
Based on this belief, we designed neural networks to learn useful
visual features by solving geo-calibration tasks. Our work provide new
ways for automatic feature learning.

### 2018-08-20 dissertation.tex
a check from the abstract to main contributions

\abstract
{ \SingleSpacing
The ultimate goal of image understanding is to transfer visual
images into numerical or symbolic descriptions of the scene
that are helpful for decision making.
%
%In general, understanding images is not trivial for machines.
%Under different scenarios, researchers divide image
%understanding into different computer vision tasks. For example,
%pedestrian detection is essential to autonomous driving, and image
%retrieval to an image searching engine.
%
Knowing when, where, and in which direction picture was taken, the
task of geo-calibration makes it possible to use imagery to understand
the world and how it changes in time. Current models for
geo-calibration are mostly deterministic, which fails to model the
the inherent uncertainties when the scene of the image is ambiguous.
Furthermore, the outputs of deterministic system can not be used by
algorithms that take probabilistic inputs. To address these
limitations, we propose a probabilistic model for camera
geo-calibration using deep neural networks.
%
While our primary contribution is geo-calibration, We also show that
learning to geo-calibrate a camera is helpful for high-level feature
extraction.
%
%Our discoveries can provide researchers new insights for both camera
%geo-calibration and DNN training.
}
\keywords{computer vision, geo-calibration, deep neural networks}

>>>>

\abstract
{ \SingleSpacing
The ultimate goal of image understanding is to transfer visual
images into numerical or symbolic descriptions of the scene
that are helpful for decision making.
%
%In general, understanding images is not trivial for machines.
%Under different scenarios, researchers divide image
%understanding into different computer vision tasks. For example,
%pedestrian detection is essential to autonomous driving, and image
%retrieval to an image searching engine.
%
Knowing when, where, and in which direction picture was taken, the
task of geo-calibration makes it possible to use imagery to understand
the world and how it changes in time. Current models for
geo-calibration are mostly deterministic, which in many cases fails to
model the inherent uncertainties when the scene of the image is
ambiguous.  Furthermore, the outputs of deterministic system can not
be used by algorithms that take probabilistic inputs. To address these
limitations, we propose a probabilistic model for camera
geo-calibration using deep neural networks.
%
While our primary contribution is geo-calibration, We also show that
learning to geo-calibrate a camera is helpful for high-level feature
extraction.
%
%Our discoveries can provide researchers new insights for both camera
%geo-calibration and DNN training.
}
\keywords{computer vision, geo-calibration, deep neural networks}
